### 2016-12-08 Paper_v1.tex
Dec8

Gene knockouts in the electron transport chain significantly reduced the performance of the CFPS reaction (Fig.~\ref{fig:oxphos_ko}).
A key finding of both the CAT and overall system state sensitivity analysis was the importance of oxidative phosphorylation.
To investigate this further, we knocked out key oxidative phosphorylation reactions in the ensemble of kinetic models to examine the effect on glucose uptake and CAT production.
%A single \textit{cyd} knockout reduced the CAT carbon yield from 7.9\% to 2.6\% (Table~\ref{tbl:yield_breakdown}).
A single \textit{cyd} knockout reduced the CAT carbon yield from 2.7\% to 0.9\% (Table~\ref{tbl:yield_breakdown}).
In addition, the glucose uptake rate was reduced compared to that of the control (no knockouts).
%On the other hand, a \textit{nuo} knockout showed a less dramatic decrease in yield,
%reducing the CAT carbon yield to 6.9\%; however, the glucose uptake rate remained similar to that of the control.
On the other hand, a \textit{nuo} knockout showed a less dramatic decrease in yield,
reducing the CAT carbon yield to 2.3\%; however, the glucose uptake rate remained similar to that of the control.
%Knocking out \textit{app} increased CAT yield to 8.3\%, but this increase was not statistically significantly different from that of the control.
Knocking out \textit{app} did not change the CAT yield (it remained at 2.7\%).
%Lastly, knocking out all three reactions reduced the CAT yield to 3.0\%, but this was not statistically significantly different from the \textit{cyd} knockout alone.
Lastly, knocking out all three reactions reduced the CAT yield to 0.7\%, similar to knocking out the \textit{cyd} alone.
Thus, the model suggested the key step in oxidative phosphorylation was catalyzed by the gene product of \textit{cyd}.
However, while disruption of \textit{cyd} significantly reduced the CAT carbon yield, it did not eliminate the ability of CFPS reaction to produce CAT.
This suggested there was a mixture of energy sources supporting CAT production, with the most significant being oxidative phosphorylation.
%The \textit{cyd} enzyme is responsible for transferring electrons with oxygen as an electron acceptor allowing the release of energy to form ATP.
%The sensitivity analysis showed oxidative phosphorylation was significant for CAT production as well as the system state.
A similar distribution of the carbon contribution to CAT yield was seen across the best-fit set and all knockouts.
In all cases, glutamate was responsible for about two-thirds of carbon consumption toward CAT.
This was due to its role in generating other amino acids and as a possible energy source for metabolism.
The much greater consumption of this one amino acid was made possible by its much larger initial condition, as it was present in the cell media in the form of magnesium glutamate, ammonium glutamate, and potassium glutamate [CITE? THESIS NOT CURRENTLY IN BIBLIOGRAPHY].
Glucose made the next largest contribution, between 20\% and 30\%, dwarfing all amino acids other than glutamate.
This makes sense, as glucose powers the entire metabolism during the first hour, including energy species synthesis and amino acid synthesis.
The knockouts in oxidative phosphorylation also reduced the energy efficiency of CAT production  (Fig.~\ref{fig:oxphos_ko}).
The best-fit set, $\Delta$app, and experimental dataset had efficiencies of 2.6\%, $\Delta$nuo had an efficiency of 2.2\%, and $\Delta$cyd and $\Delta$cyd$\Delta$nuo$\Delta$app had efficiencies of 0.8\%.
This was in line with the effect of knockouts on yield, which in turn was more or less proportional to the total amount of CAT produced in each case.
Taken together, these results show that oxidative phosphorylation, particularly \textit{cyd}, is important to efficient CAT production.

Furthermore, we examined the proportion of amino acid consumption that was due to CAT synthesis in the best-fit set (Table~\ref{tbl:AA_breakdown}).
Eight of the 20 amino acids had a percentage of 100\% or higher.
For six of these (isoleucine, leucine, methionine, phenylalanine, tryptophan, valine), synthesis was inactive or negligible, meaning that CAT synthesis was the only reaction that consumed these species.
This is why methionine, and to a lesser extent isoleucine, leucine, and valine do not fit the experimental data as well as some other amino acids; consumption via CAT synthesis was not enough to fit the data, and no other reactions exist in the network to consume these species.
For histidine and tyrosine, synthesis was active, allowing the percentage to be higher than 100\%.
Another ten amino acids registered percentages lower than 100\%, meaning that they were consumed by other pathways in the network.
The most extreme of these is glutamate, with a percentage of 0.2\%, because it was consumed by numerous reactions for the synthesis of other amino acids and as an energy substrate.
Likewise, arginine, asparagine, glycine, and proline are consumed in the production of other amino acids, while cysteine and lysine are consumed to form other metabolites.
%ARG R_arg_deg1-4 go to GLU
%ASN R_asn_deg goes to ASP
%ASP R_thr, maybe negligible
%CYS cys_deg goes to pyr
%GLY R_gar goes to GLU
%LYS lys_deg goes to cadav
%PRO pro_deg goes to GLU
%SER R_glyA, maybe negligible
%THR R_thr_deg4, maybe negligible
Finally, alanine and glutamine increased during the timecourse, meaning that synthesis outweighed all consumption (including consumption toward CAT).

Sequence-specific flux balance analysis (ssFBA) predicted optimal CAT yields with no adjustable parameters (Fig.~\ref{fig:CATProdYield}).
Before exploring CFPS optimality, we first validated the ssFBA approach by comparing simulated and measured concentrations of CAT for the first hour of glucose consumption.
We chose this time window (during the first phase of CAT production) because it was approximately linear in both glucose consumption and by-/production formation.
The ssFBA calculation had no adjustable parameters; bounds on transcription and translation rates, and biochemical fluxes were either estimated from data, or
from mechanistic models parameterized from literature.
Uncertainty in experimental factors such as RNA polymerase, ribosome concentrations, elongation rates, or the upper bounds for oxygen and glucose consumption rates was addressed
by sampling plausible ranges for these parameters.
The ensemble of ssFBA simulations predicted CAT formation as a function of time during the first hour of production when constrained by the experimental metabolite data (Fig.~\ref{fig:CATProdYield}A).
Thus, the molecular description of transcription and translation were consistent with experimental measurements.
Next, to gauge the performance of the CFPS reaction, we next calculated the CAT carbon yield for three classes of constraints:
(i) theoretical maximum glucose, amino acid and oxygen upper bounds, and realistic transcriptional/translational constraints;
(ii) theoretical maximum glucose, amino acid and oxygen upper bounds, realistic transcriptional/translational constraints and knockouts of amino acid synthesis reactions of amino acids supplemented in the \textit{E. coli} extract preparation.  
(iii) metabolite fluxes constrained by the CAT data, and realistic transcriptional/translational constraints and knockouts of amino acid synthesis reactions of amino acids supplemented in the \textit{E. coli} extract preparation (Fig.~\ref{fig:CATProdYield}B).
The physiological theoretical maximum CAT carbon yield (case i) was 21.8\% $\pm$ 2.8\% (Fig.~\ref{fig:CATProdYield}B, left); this represents optimal network performance if glucose, oxygen and amino acids were produced or consumed at their upper bounds, with bounded transcription and translation rates (96\% without glucose contribution in the carbon yield calculation).
For case ii, the optimal CAT carbon yield decreased to 18.2\% $\pm$ 3.0\% (Fig.~\ref{fig:CATProdYield}B, middle).
Lastly, when metabolite constraints were applied with experimental measurements of measurements (case iii), the predicted carbon yield was 6.0\% $\pm$ 2.0\%
%By comparison, the best-fit parameter set for the kinetic model predicted a CAT carbon yield of 8.6\% $\pm$ 0.4\%, equivalent to 23\% of the theoretical maximum (i) and 35\% of the physiological case (ii).
By comparison, the best-fit parameter set for the kinetic model predicted a CAT carbon yield (without arginine and glutamate) of 7.9\%, equivalent to 36\% of the optimal case (i).
The experimental dataset had a CAT carbon yield of 8.2\%, similar to both the kinetic model and the experimentally constrained ssFBA calculation (case iii).
However, the energy efficiency for case i was 18.1\% and 15.5\% for the second case, while for the experimentally constrained case it was 3.9\%.  
Thus, while the CFPS reaction was not optimal, the ssFBA calculations suggested that an approximately three-fold increase in carbon yield was theoretically possible.

@@

\subsection*{Sequence specific calculation of carbon yield.}
We estimated the theoretical maximum CAT carbon yield using sequence-specific flux balance analysis (ssFBA) \cite{Allen:2003aa}.
The sequence specific flux balance analysis problem was formulated as a linear program:
\begin{equation}
 \begin{multlined}
	\qquad \qquad \qquad \max_{\boldsymbol{w}}{} \! \left( w_{TL} = \mathbf{\boldsymbol{\theta}}^T \boldsymbol{w} \right) \\
	\mathrm{Subject \; to:}
	 \; \; \mathbf{S}\mathbf{w}=\mathbf{0} \\
\alpha_i \leq w_i \leq \beta_i  \qquad i=1,2,\hdots,\mathcal{R}
 \end{multlined}
\end{equation}
where $\mathbf{S}$ denotes the stoichiometric matrix, $\mathbf{w}$ denotes the unknown flux vector, $\boldsymbol{\theta}$ denotes the objective selection vector
and $\alpha_i$ and $\beta_i$ denote the lower and upper bounds on flux $w_{i}$, respectively.
The stoichiometry of the kinetic model was used for the ssFBA calculations, with the execpetion of the transcription and translation rates.
The transcription (TX) and translation (TL) stoichiometry was modeled using the template reactions taken from Allen and Palsson \cite{Allen:2003aa}:
\begin{eqnarray*}
\mathrm{{G}_{\mathcal{P}}}+\mathrm{{R}_{1}} &\longrightarrow& \mathrm{G_{\mathcal{P}}^{*}} \\
\mathrm{G_{\mathcal{P}}^{*}}+\sum_{\mathrm{k\in\left\{A,C,G,U\right\}}}\eta_{k}\cdot \mathrm{\left\{k\right\}TP} &\overset{\rm TX}{\longrightarrow}& \mathrm{mRNA}+\mathrm{G_{\mathcal{P}}}+\mathrm{R_{1}}+ \mathrm{\sum_{k\in\left\{A,C,G,U\right\}}2\eta_{k}\cdot Pi}\\
\mathrm{mRNA} &\longrightarrow& \mathrm{\sum_{k\in\left\{A,C,G,U\right\}}\eta_{k}\cdot \left\{k\right\}MP} \\
\mathrm{mRNA}+\mathrm{R_{2}} &\longrightarrow& \mathrm{R_{2}^{*}} \\
\mathrm{\alpha_{j}\cdot AA_{j}+\alpha_{j}\cdot tRNA+\alpha_{j}\cdot ATP} &\longrightarrow& \mathrm{\alpha_{j}\cdot AA_{j}-tRNA_{j}+}\\
&& \mathrm{\qquad \alpha_{j}\cdot AMP+2\alpha_{j}\cdot Pi} \qquad{j=1,2,\hdots,20}\\
\mathrm{R_{2}^{*}+\sum_{j\in\left\{AA\right\}}\alpha_{j}\cdot \Big(AA_{j}-tRNA_{j}+2\cdot GTP\Big)} &\stackrel{\rm TL}{\longrightarrow}& \mathcal{P}+\mathrm{R_{2}+mRNA}+\\
&& \mathrm{\quad +\sum_{j\in\left\{AA\right\}}\alpha_{j}\cdot\Big(tRNA+2\cdot GDP+2\cdot Pi\Big)}
\end{eqnarray*}
where $G_{\mathcal{P}}$ denotes the gene encoding protein product $\mathcal{P}$,
$\rm R_{1}$ denotes the concentration of RNA polymerase,
$G_{\mathcal{P}}^{*}$ denotes the gene bounded by the RNA polymerase,
$\eta_{i}$ and $ \alpha_{j}$ denote the stoichiometric coefficients for nucleotide and amino acid, respectively,
$\rm Pi$ denotes inorganic phosphate,
$\rm R_{2}$ denotes the ribosome concentration,
$\rm R_{2}^{*}$ denotes bounded ribosome,
and $AA_{j}$ denotes $j^{th}$ amino acid.

@@

The objective of the sequence specific flux balance calculation was to maximize the rate of CAT translation, $w_{TL}$.
The total glucose uptake rate was bounded by [0,40 mM/h] according to experimental data;
while the amino acid uptake rates were bounded by [0,30 mM/h], but did not reach the maximum flux.
The CAT gene and protein sequences were taken from literature.
The sequence specific flux balance linear program was solved using the GNU Linear Programming Kit (GLPK) v4.52 \cite{GLPK}.

@@

\subsection*{Calculation of the carbon yield.}
The CAT carbon yield ($Y_{C}^{CAT}$) was calculated as the ratio of carbon produced as CAT divided by the carbon consumed as reactants (glucose and amino acids):
\begin{equation}\label{eqn:yield-definition}
	Y_{C}^{CAT}=\frac{\Delta\texttt{CAT}\cdot C_{CAT}}{\displaystyle\sum_{i=1}^{\mathcal{R}}\max(\Delta m_{i},0)\cdot C_{m_i}}
\end{equation}
where $\Delta\texttt{CAT}$ denotes the abundance of CAT produced, $C_{CAT}$ denotes carbon number of CAT, $\mathcal{R}$ denotes the number of reactants,
$\Delta m_{i}$ denotes the amount of the $i^{th}$ reactant consumed (never allowed to be negative), and $C_{m_i}$ denotes the carbon number of the $i^{th}$ reactant.
Arginine and glutamate were excluded from the yield calculations for the experimental yield calculation, as no experimental measurements were available for these amino acids.
Yield of the best-fit parameter set and the experimental data were calculated by setting $\Delta\texttt{CAT}$ equal to the final minus the initial CAT concentration,
and setting $\Delta m_{i}$ equal to the initial minus the final reactant concentration.
The individual CAT production and substrate consumption terms for the best-fit set, kinetic models with knockouts, and experimental data are shown in Table~\ref{tbl:yield_breakdown}.
Consumption of amino acids via CAT synthesis was calculated as a percentage of total net consumption for the best-fit set (Table~\ref{tbl:AA_breakdown}).
Consumption toward CAT was calculated as CAT production times the stoichiometric coefficient for that amino acid in the CAT synthesis reaction.
Total net consumption was calculated as amino acid concentration at 3 hours minus concentration at 0 hours, and was excluded if negative.

\subsection*{Calculation of energy efficiency.}
Energy efficiency was also calculated for best-fit set, knockouts, and experimental data, formulated as:
\begin{equation}\label{eqn:energy-efficiency-definition}
	\rm Efficiency=\frac{\Delta\texttt{CAT}\cdot \left(2\cdot\rm {ATP}_{CAT}+\rm {GTP}_{CAT}\right)}{\Delta\texttt{GLC}\cdot \rm {ATP}_{GLC}}
\end{equation}
where $\rm Efficiency$ denotes the energy efficiency of CAT production, $\rm {ATP}_{CAT}$ denotes the stoichiometric coefficient of ATP in CAT synthesis (multiplied by 2 because AMP, rather than ADP, is a product of CAT synthesis), $\rm {GTP}_{CAT}$ denotes the stoichiometric coefficient of GTP in CAT synthesis, and $\rm {ATP}_{GLC}$ denotes the number of ATP molecules ideally produced from one molecule of glucose.
$\rm {ATP}_{CAT}$ = 219, $\rm {GTP}_{CAT}$ = 438, $\rm {ATP}_{GLC}$ = 17.

Energy efficiency was also calculated for sequence specific flux balance analysis, formulated as:
\begin{equation}\label{eqn:energy-efficiency-definition}
	\rm Efficiency=\frac{\Delta\texttt{CAT}\cdot \left(2\left({ATP}_{TX}+\rm {CTP}_{TX}+\rm {GTP}_{TX}+\rm {UTP}_{TX}\right)+\rm2\cdot\rm {ATP}_{TL}+\rm {GTP}_{TL}\right)}{\Delta\texttt{GLC}\cdot \rm {ATP}_{GLC}}
\end{equation}
where $\rm Efficiency$ denotes the energy efficiency of CAT production, $\rm {ATP}_{TX}$, $\rm {CTP}_{TX}$, $\rm {GTP}_{TX}$, $\rm {UTP}_{TX}$ denote the stoichiometric coefficient of each energy species for CAT transcription, $\rm {ATP}_{TL}$ and $\rm {GTP}_{TL}$ denote the stoichiometric coefficient of ATP and GTP, respectively, in CAT translation, and $\rm {ATP}_{GLC}$ denotes the number of ATP molecules produced from one molecule of glucose.
$\rm {ATP}_{TX}$ = 176, $\rm {CTP}_{TX}$ = 144, $\rm {GTP}_{TX}$ = 151, $\rm {UTP}_{TX}$ = 189, $\rm {ATP}_{TL}$ = 219, $\rm {GTP}_{TL}$ = 438, $\rm {ATP}_{GLC}$ = 17.
%The flux balance analysis problem was solved using the GNU Linear Programming Kit (v4.52)

% The solution flux vector was used to calculate the carbon yield of CAT for the three FBA cases.
% Glucose, oxygen, and amino acids were modeled as being imported into the system, while CAT synthesis and metabolite byproduct formation were modeled as export from the system.
%Need to talk more about biochemical benefits and importance for biochemical problems

>>>>

Gene knockouts in the electron transport chain significantly reduced the performance of the CFPS reaction (Fig.~\ref{fig:oxphos_ko}).
A key finding of both the CAT and overall system state sensitivity analysis was the importance of oxidative phosphorylation.
To investigate this further, we knocked out key oxidative phosphorylation reactions in the ensemble of kinetic models to examine the effect on glucose uptake and CAT production.
%A single \textit{cyd} knockout reduced the CAT carbon yield from 7.9\% to 2.6\% (Table~\ref{tbl:yield_breakdown}).
A single \textit{cyd} knockout reduced the CAT carbon yield from 2.7\% to 0.9\% (Table~\ref{tbl:yield_breakdown}).
In addition, the glucose uptake rate was reduced compared to that of the control (no knockouts).
%On the other hand, a \textit{nuo} knockout showed a less dramatic decrease in yield,
%reducing the CAT carbon yield to 6.9\%; however, the glucose uptake rate remained similar to that of the control.
On the other hand, a \textit{nuo} knockout showed a less dramatic decrease in yield,
reducing the CAT carbon yield to 2.3\%; however, the glucose uptake rate remained similar to that of the control.
%Knocking out \textit{app} increased CAT yield to 8.3\%, but this increase was not statistically significantly different from that of the control.
Knocking out \textit{app} did not change the CAT yield (it remained at 2.7\%).
%Lastly, knocking out all three reactions reduced the CAT yield to 3.0\%, but this was not statistically significantly different from the \textit{cyd} knockout alone.
Lastly, knocking out all three reactions reduced the CAT yield to 0.7\%, similar to knocking out the \textit{cyd} alone.
Thus, the model suggested the key step in oxidative phosphorylation was catalyzed by the gene product of \textit{cyd}.
However, while disruption of \textit{cyd} significantly reduced the CAT carbon yield, it did not eliminate the ability of CFPS reaction to produce CAT.
This suggested there was a mixture of energy sources supporting CAT production, with the most significant being oxidative phosphorylation.
%The \textit{cyd} enzyme is responsible for transferring electrons with oxygen as an electron acceptor allowing the release of energy to form ATP.
%The sensitivity analysis showed oxidative phosphorylation was significant for CAT production as well as the system state.
A similar distribution of the carbon contribution to CAT yield was seen across the best-fit set and all knockouts.
In all cases, glutamate was responsible for about two-thirds of carbon consumption toward CAT.
This was due to its role in generating other amino acids and as a possible energy source for metabolism.
The much greater consumption of this one amino acid was made possible by its much larger initial condition, as it was present in the cell media in the form of magnesium glutamate, ammonium glutamate, and potassium glutamate [CITE? THESIS NOT CURRENTLY IN BIBLIOGRAPHY].
Glucose made the next largest contribution, between 20\% and 30\%, dwarfing all amino acids other than glutamate.
This makes sense, as glucose powers the entire metabolism during the first hour, including energy species synthesis and amino acid synthesis.
%We also investigated the energy efficiency of CAT production across the kinetic models and experimental dataset.
%The best-fit set, $\Delta$app, and experimental data had energy efficiencies of 2.6\%, $\Delta$nuo had an efficiency of 2.2\%, and $\Delta$cyd and $\Delta$cyd$\Delta$nuo$\Delta$app had efficiencies of 0.8\%.
%The best-fit set, $\Delta$app, and experimental data had energy efficiencies of 2.6\%, calculated as a ratio of CAT production to glucose consumption, each converted into an equivalent number of ATP molecules.
%$\Delta$nuo had an efficiency of 2.2\%, and $\Delta$cyd and $\Delta$cyd$\Delta$nuo$\Delta$app had efficiencies of 0.8\%.
%The energy efficiencies across the different cases were more or less proportional to the carbon yields, and in turn the total amount of CAT synthesis.
Taken together, these results show that oxidative phosphorylation, particularly \textit{cyd}, is important to efficient CAT production.

%Furthermore, we examined the proportion of amino acid consumption that was due to CAT synthesis in the best-fit set (Table~\ref{tbl:AA_breakdown}).
To better understand where carbon flux goes besides toward CAT, we examined the proportion of amino acid consumption that was due to CAT synthesis in the best-fit set (Table~\ref{tbl:AA_breakdown}).
Eight of the 20 amino acids had a flux-toward-CAT percentage of 100\% or higher.
For six of these (isoleucine, leucine, methionine, phenylalanine, tryptophan, valine), synthesis was inactive or negligible, meaning that CAT synthesis was the only reaction that consumed these species.
This is why methionine, and to a lesser extent isoleucine, leucine, and valine do not fit the experimental data as well as some other amino acids; consumption via CAT synthesis was not enough to fit the data, and no other reactions exist in the network to consume these species.
For histidine and tyrosine, synthesis was active, allowing the percentage to be higher than 100\%.
Another ten amino acids registered percentages lower than 100\%, meaning that they were consumed by other pathways in the network.
The most extreme of these is glutamate, with a percentage of 0.2\%, because it was consumed by numerous reactions for the synthesis of other amino acids and as an energy substrate.
Likewise, arginine, asparagine, glycine, and proline are consumed in the production of other amino acids, while cysteine and lysine are consumed to form other metabolites.
%ARG R_arg_deg1-4 go to GLU
%ASN R_asn_deg goes to ASP
%ASP R_thr, maybe negligible
%CYS cys_deg goes to pyr
%GLY R_gar goes to GLU
%LYS lys_deg goes to cadav
%PRO pro_deg goes to GLU
%SER R_glyA, maybe negligible
%THR R_thr_deg4, maybe negligible
Finally, alanine and glutamine increased during the timecourse, meaning that synthesis outweighed all consumption (including consumption toward CAT).

Sequence-specific flux balance analysis (ssFBA) predicted optimal CAT yields with no adjustable parameters (Fig.~\ref{fig:CATProdYield}).
Before exploring CFPS optimality, we first validated the ssFBA approach by comparing simulated and measured concentrations of CAT for the first hour of glucose consumption.
We chose this time window (during the first phase of CAT production) because it was approximately linear in both glucose consumption and by-/production formation.
The ssFBA calculation had no adjustable parameters; bounds on transcription and translation rates, and biochemical fluxes were either estimated from data, or
from mechanistic models parameterized from literature.
Uncertainty in experimental factors such as RNA polymerase, ribosome concentrations, elongation rates, or the upper bounds for oxygen and glucose consumption rates was addressed
by sampling plausible ranges for these parameters.
The ensemble of ssFBA simulations predicted CAT formation as a function of time during the first hour of production when constrained by the experimental metabolite data (Fig.~\ref{fig:CATProdYield}A).
Thus, the molecular description of transcription and translation were consistent with experimental measurements.
Next, to gauge the performance of the CFPS reaction, we next calculated the CAT carbon yield for three classes of constraints:
(i) theoretical maximum glucose, amino acid and oxygen upper bounds, and realistic transcriptional/translational constraints;
(ii) theoretical maximum glucose, amino acid and oxygen upper bounds, realistic transcriptional/translational constraints and knockouts of amino acid synthesis reactions of amino acids supplemented in the \textit{E. coli} extract preparation.  
(iii) metabolite fluxes constrained by the CAT data, and realistic transcriptional/translational constraints and knockouts of amino acid synthesis reactions of amino acids supplemented in the \textit{E. coli} extract preparation (Fig.~\ref{fig:CATProdYield}B).
The physiological theoretical maximum CAT carbon yield (case i) was 21.8\% $\pm$ 2.8\% (Fig.~\ref{fig:CATProdYield}B, left); this represents optimal network performance if glucose, oxygen and amino acids were produced or consumed at their upper bounds, with bounded transcription and translation rates (96\% without glucose contribution in the carbon yield calculation).
For case ii, the optimal CAT carbon yield decreased to 18.2\% $\pm$ 3.0\% (Fig.~\ref{fig:CATProdYield}B, middle).
Lastly, when metabolite constraints were applied with experimental measurements of measurements (case iii), the predicted carbon yield was 6.0\% $\pm$ 2.0\%.
By comparison, the best-fit parameter set for the kinetic model predicted a CAT carbon yield (without arginine and glutamate) of 7.9\%, equivalent to 36\% of the optimal case (i).
The experimental dataset had a CAT carbon yield of 8.2\%, similar to both the kinetic model and the experimentally constrained ssFBA calculation (case iii).
We also investigated the energy efficiency of CAT production for the best-fit set and the three sequence-specific cases, based on the amount of CAT production versus ATP production.
For cases i and ii the energy efficiencies were 99.8\% and 94.4\%, respectively, while for case iii it was 31.9\%.
This dramatic decrease in efficiency when fluxes are constrained to data makes sense, as the network is forced toward a multitude of pathways that may not contribute to CAT production.
However, the case iii efficiency was still much higher than that of the best-fit kinetic model (9.2\%).
This is likely because of the steady-state assumption and the ability to choose the optimum from a variety of flux distributions; meanwhile, the kinetic model diverts flux toward the accumulation of sub-optimal metabolites.
Thus, the CFPS reaction was not optimal; the ssFBA calculations suggested that an approximately three-fold increase in carbon yield and an eleven-fold increase in energy efficiency were theoretically possible.

@@

\subsection*{Sequence-specific calculation of carbon yield.}
We estimated the theoretical maximum CAT carbon yield using sequence-specific flux balance analysis (ssFBA) \cite{Allen:2003aa}.
The sequence-specific flux balance analysis problem was formulated as a linear program:
\begin{equation}
 \begin{multlined}
	\qquad \qquad \qquad \max_{\boldsymbol{w}}{} \! \left( w_{TL} = \mathbf{\boldsymbol{\theta}}^T \boldsymbol{w} \right) \\
	\mathrm{Subject \; to:}
	 \; \; \mathbf{S}\mathbf{w}=\mathbf{0} \\
\alpha_i \leq w_i \leq \beta_i  \qquad i=1,2,\hdots,\mathcal{R}
 \end{multlined}
\end{equation}
where $\mathbf{S}$ denotes the stoichiometric matrix, $\mathbf{w}$ denotes the unknown flux vector, $\boldsymbol{\theta}$ denotes the objective selection vector
and $\alpha_i$ and $\beta_i$ denote the lower and upper bounds on flux $w_{i}$, respectively.
The stoichiometry of the kinetic model was used for the ssFBA calculations, with the execpetion of the transcription and translation rates.
The transcription (TX) and translation (TL) stoichiometry was modeled using the template reactions taken from Allen and Palsson \cite{Allen:2003aa}:
\begin{eqnarray*}
\mathrm{{G}_{\mathcal{P}}}+\mathrm{{R}_{1}} &\longrightarrow& \mathrm{G_{\mathcal{P}}^{*}} \\
\mathrm{G_{\mathcal{P}}^{*}}+\sum_{\mathrm{k\in\left\{A,C,G,U\right\}}}\eta_{k}\cdot \mathrm{\left\{k\right\}TP} &\overset{\rm TX}{\longrightarrow}& \mathrm{mRNA}+\mathrm{G_{\mathcal{P}}}+\mathrm{R_{1}}+ \mathrm{\sum_{k\in\left\{A,C,G,U\right\}}2\eta_{k}\cdot Pi}\\
\mathrm{mRNA} &\longrightarrow& \mathrm{\sum_{k\in\left\{A,C,G,U\right\}}\eta_{k}\cdot \left\{k\right\}MP} \\
\mathrm{mRNA}+\mathrm{R_{2}} &\longrightarrow& \mathrm{R_{2}^{*}} \\
\mathrm{\alpha_{j}\cdot AA_{j}+\alpha_{j}\cdot tRNA+\alpha_{j}\cdot ATP} &\longrightarrow& \mathrm{\alpha_{j}\cdot AA_{j}-tRNA_{j}+}\\
&& \mathrm{\qquad \alpha_{j}\cdot AMP+2\alpha_{j}\cdot Pi} \qquad{j=1,2,\hdots,20}\\
\mathrm{R_{2}^{*}+\sum_{j\in\left\{AA\right\}}\alpha_{j}\cdot \Big(AA_{j}-tRNA_{j}+2\cdot GTP\Big)} &\stackrel{\rm TL}{\longrightarrow}& \mathcal{P}+\mathrm{R_{2}+mRNA}+\\
&& \mathrm{\quad +\sum_{j\in\left\{AA\right\}}\alpha_{j}\cdot\Big(tRNA+2\cdot GDP+2\cdot Pi\Big)}
\end{eqnarray*}
where $G_{\mathcal{P}}$ denotes the gene encoding protein product $\mathcal{P}$,
$\rm R_{1}$ denotes the concentration of RNA polymerase,
$G_{\mathcal{P}}^{*}$ denotes the gene bounded by the RNA polymerase,
$\eta_{i}$ and $ \alpha_{j}$ denote the stoichiometric coefficients for nucleotide and amino acid, respectively,
$\rm Pi$ denotes inorganic phosphate,
$\rm R_{2}$ denotes the ribosome concentration,
$\rm R_{2}^{*}$ denotes bounded ribosome,
and $AA_{j}$ denotes $j^{th}$ amino acid.

@@

The objective of the sequence-specific flux balance calculation was to maximize the rate of CAT translation, $w_{TL}$.
The total glucose uptake rate was bounded by [0,40 mM/h] according to experimental data;
while the amino acid uptake rates were bounded by [0,30 mM/h], but did not reach the maximum flux.
The CAT gene and protein sequences were taken from literature.
The sequence-specific flux balance linear program was solved using the GNU Linear Programming Kit (GLPK) v4.52 \cite{GLPK}.

@@

\subsection*{Calculation of the carbon yield.}
The CAT carbon yield ($Y_{C}^{CAT}$) was calculated as the ratio of carbon produced as CAT divided by the carbon consumed as reactants (glucose and amino acids):
\begin{equation}\label{eqn:yield-definition}
	Y_{C}^{CAT}=\frac{\Delta\texttt{CAT}\cdot C_{CAT}}{\displaystyle\sum_{i=1}^{\mathcal{R}}\max(\Delta m_{i},0)\cdot C_{m_i}}
\end{equation}
where $\Delta\texttt{CAT}$ denotes the abundance of CAT produced, $C_{CAT}$ denotes carbon number of CAT, $\mathcal{R}$ denotes the number of reactants,
$\Delta m_{i}$ denotes the amount of the $i^{th}$ reactant consumed (never allowed to be negative), and $C_{m_i}$ denotes the carbon number of the $i^{th}$ reactant.
Arginine and glutamate were excluded from the yield calculations for the experimental yield calculation, as no experimental measurements were available for these amino acids.
Yield of the best-fit parameter set and the experimental data were calculated by setting $\Delta\texttt{CAT}$ equal to the final minus the initial CAT concentration,
and setting $\Delta m_{i}$ equal to the initial minus the final reactant concentration.
The individual CAT production and substrate consumption terms for the best-fit set, kinetic models with knockouts, and experimental data are shown in Table~\ref{tbl:yield_breakdown}.
Total net consumption of amino acids and amino acid consumption via CAT synthesis were calculated for the best-fit set (Table~\ref{tbl:AA_breakdown}).
Total net consumption was calculated as amino acid concentration at 0 hours minus concentration at 3 hours; it was negative if synthesis outweighed consumption.
Consumption toward CAT was calculated as CAT concentration at 3 hours minus concentration at 0 hours, times the stoichiometric coefficient for that amino acid in the CAT synthesis reaction.
The difference between these was defined as other consumption, equal to consumption from reactions other than CAT synthesis minus amino acid production.

\subsection*{Calculation of energy efficiency.}
%Energy efficiency was also calculated for best-fit set, knockouts, and experimental data, formulated as:
%\begin{equation}\label{eqn:energy-efficiency-definition}
%	\rm Efficiency=\frac{\Delta\texttt{CAT}\cdot \left(2\cdot\rm {ATP}_{CAT}+\rm {GTP}_{CAT}\right)}{\Delta\texttt{GLC}\cdot \rm {ATP}_{GLC}}
%\end{equation}
%where $\rm Efficiency$ denotes the energy efficiency of CAT production, $\rm {ATP}_{CAT}$ denotes the stoichiometric coefficient of ATP in CAT synthesis (multiplied by 2 because AMP, rather than ADP, is a product of CAT synthesis), $\rm {GTP}_{CAT}$ denotes the stoichiometric coefficient of GTP in CAT synthesis, and $\rm {ATP}_{GLC}$ denotes the number of ATP molecules ideally produced from one molecule of glucose.
%$\rm {ATP}_{CAT}$ = 219, $\rm {GTP}_{CAT}$ = 438, $\rm {ATP}_{GLC}$ = 17.

Energy efficiency was also calculated for the best-fit set:
\begin{equation}\label{eqn:energy-efficiency-definition}
	\rm Efficiency=\frac{\Delta\texttt{CAT}\cdot \left(2\cdot\rm {ATP}_{CAT}+\rm {GTP}_{CAT}\right)}{\Delta\texttt{ATP}}
\end{equation}
where $\rm Efficiency$ denotes the energy efficiency of CAT production, $\rm {ATP}_{CAT}$ denotes the stoichiometric coefficient of ATP in CAT synthesis (multiplied by 2 because AMP, rather than ADP, is a product of CAT synthesis), $\rm {GTP}_{CAT}$ denotes the stoichiometric coefficient of GTP in CAT synthesis, and $\rm \Delta ATP$ denotes the amount of ATP production in 3 hours, equal to the sum of ATP-producing fluxes integrated over the timecourse.
$\rm {ATP}_{CAT}$ = 219, $\rm {GTP}_{CAT}$ = 438.
%, $\rm \Delta CAT$ = 0.018 mM/h, $\rm \Delta ATP$ = 170 mM/h.

%Energy efficiency was also calculated for sequence-specific flux balance analysis, formulated as:
%\begin{equation}\label{eqn:energy-efficiency-definition}
%	\rm Efficiency=\frac{\Delta\texttt{CAT}\cdot \left(2\left({ATP}_{TX}+\rm {CTP}_{TX}+\rm {GTP}_{TX}+\rm {UTP}_{TX}\right)+\rm2\cdot\rm {ATP}_{TL}+\rm {GTP}_{TL}\right)}{\Delta\texttt{GLC}\cdot \rm {ATP}_{GLC}}
%\end{equation}
%where $\rm Efficiency$ denotes the energy efficiency of CAT production, $\rm {ATP}_{TX}$, $\rm {CTP}_{TX}$, $\rm {GTP}_{TX}$, $\rm {UTP}_{TX}$ denote the stoichiometric coefficient of each energy species for CAT transcription, $\rm {ATP}_{TL}$ and $\rm {GTP}_{TL}$ denote the stoichiometric coefficient of ATP and GTP, respectively, in CAT translation, and $\rm {ATP}_{GLC}$ denotes the number of ATP molecules produced from one molecule of glucose.
%$\rm {ATP}_{TX}$ = 176, $\rm {CTP}_{TX}$ = 144, $\rm {GTP}_{TX}$ = 151, $\rm {UTP}_{TX}$ = 189, $\rm {ATP}_{TL}$ = 219, $\rm {GTP}_{TL}$ = 438, $\rm {ATP}_{GLC}$ = 17.
